Intentional release of native species undermines ecological stability

### 2022-10-03 rmd/supporting_info.Rmd
Update supporting_info.Rmd

Under each simulation scenario, we ran `r attr(list_param[[1]], "n_sim")` timesteps of community dynamics for $100$ values of release level $R$ ($0$ to $500$ with an equal interval). We introduced $50$ individuals of each species and allowed them to grow with no intentional release for the first `r unique(list_sim[[1]]$n_warmup)` timesteps (initialization). After the initialization period, we released $R$ individuals of the enhanced species every timestep over the next `r unique(list_sim[[1]]$n_burnin)` timesteps to reach a new equilibrium with the intentional release (burn-in period). We continued the simulation run with intentional release and saved the last `r unique(list_sim[[1]]$n_timestep)` timesteps, which were used to obtain the following summary statistics of the whole community $\sum_i^S N_i$: the coefficient of variation (CV), the temporal mean ($\mu$), and the SD ($\sigma$).

@@

# Supplementary text

@@

The analysis of a two-species community revealed the distinct parameter space predicting a stabilizing, neutral, or destabilizing effect of intentional release. The left column in **Figure S**\@ref(fig:fig-2sp-w)A) shows the predictions in scenarios with no environmental stochasticity ($\sigma_{\varepsilon} =$ `r min(list_param[[1]]$sd_env)`) and weak competition ($\bar{\alpha}=$ `r min(list_param[[1]]$alpha)`). For lower values of intrinsic growth rate $r_i$ ($<1.5$), intentional release had neutral (with higher $K$) to destabilizing effects (with lower $K$) on community dynamics. In the meantime, stabilization with intentional release occurred for higher values of $r_i$ ($\ge 1.5$) by dampening limit cycles or chaos (**Figure S**\@ref(fig:fig-2sp-s)), which is reflected in decreased SD $\sigma$ (**Figure S**\@ref(fig:fig-2sp-w)). However, the long-term mean of community density $\mu$ consistently decreased with intentional release in almost all cases except for only a few combinations of $r_i$ and $K$ (**Figure S**\@ref(fig:fig-2sp-w)). Environmental stochasticity did not change the qualitative patterns of deterministic simulations but greatly expanded the parameter space with destabilizing effects of intentional release (the right columns in **Figure S**\@ref(fig:fig-2sp-w)**A and B**). Increasing the competition coefficient $\bar{\alpha}$ shifted the region with the destabilizing effects to higher carrying capacities ($\bar{\alpha}=$ `r max(list_param[[1]]$alpha)`, **Figure S**\@ref(fig:fig-2sp-s)); otherwise, general patterns were similar to the weak competition scenario.

>>>>

Under each simulation scenario, we ran `r attr(list_param[[1]], "n_sim")` timesteps of community dynamics for $100$ values of release level $R$ ($0$ to $500$ with an equal interval). We introduced $50$ individuals of each species and allowed them to grow with no intentional release for the first `r unique(list_sim[[1]]$n_warmup)` timesteps (initialization). After the initialization period, we released $R$ individuals of the enhanced species every timestep over the next `r unique(list_sim[[1]]$n_burnin)` timesteps to reach a new equilibrium with the intentional release (burn-in period). We continued the simulation run with intentional release and saved the last `r unique(list_sim[[1]]$n_timestep)` timesteps, which were used to obtain the following summary statistics of the whole community $\sum_i^S N_i$ ($S$: the number of species): the coefficient of variation (CV), the temporal mean ($\mu$), and the SD ($\sigma$). The results of the two-species system were reported in [Supplementary Text].

@@

# Supplementary Text

@@

The analysis of a two-species community revealed the distinct parameter space predicting a stabilizing, neutral, or destabilizing effect of intentional release. The left column in **Figure S**\@ref(fig:fig-2sp-w)A) shows the predictions in scenarios with no environmental stochasticity ($\sigma_{\varepsilon} =$ `r min(list_param[[1]]$sd_env)`) and weak competition ($\bar{\alpha}=$ `r min(list_param[[1]]$alpha)`). For lower values of intrinsic growth rate $r_i$ ($<1.5$), intentional release had neutral (with higher $K$) to destabilizing effects (with lower $K$) on community dynamics. In the meantime, stabilization with intentional release occurred for higher values of $r_i$ ($\ge 1.5$) by dampening limit cycles or chaos (**Figure S**\@ref(fig:fig-2sp-s)), which is reflected in decreased SD $\sigma$ (**Figure S**\@ref(fig:fig-2sp-w)). However, the long-term mean of community density $\mu$ consistently decreased with intentional release in almost all cases except for only a few combinations of $r_i$ and $K$ (**Figure S**\@ref(fig:fig-2sp-w)).

Environmental stochasticity did not change the qualitative patterns of deterministic simulations; however, it greatly expanded the parameter space with destabilizing effects of intentional release (the right columns in **Figure S**\@ref(fig:fig-2sp-w)**A and B**). In particular, destabilizing effects were more broadly observed when intrinsic growth rates were low.

Increasing the competition coefficient $\bar{\alpha}$ shifted the region with the destabilizing effects to higher carrying capacities ($\bar{\alpha}=$ `r max(list_param[[1]]$alpha)`, **Figure S**\@ref(fig:fig-2sp-s)); otherwise, general patterns were similar to the weak competition scenario.
